---
title: "Correlation between GDP and Broad Money"
author: "Seiji Hirano"
date: "2022-12-18"
output: github_document
---

## Setting Up My Environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r setting up environments}
library(tidyverse)
library(ggpubr)

oecd_avgwage <- read_csv('/Users/hiranoseiji/Downloads/Data_Analysis/Macro_Economic_Wage_Analysis/oecd_avgwage.csv')

picked_country <- c('AUS', 'CAN', 'DEU', 'FRA', 'GBR', 'ITA', 'JPN', 'USA')
```

# About This Project
**Average Wages** is one of the most important indicators in economics. Many organizations target employee pay at the average rate to ensure employees receive competitive compensation while the business is still managing overall costs. 
<br>
<br>
The graph below shows changes in Average Wages over 30 years in 8 OECD countries. While Average Wages have increased from time to time in most countries, Average Wages in countries like **Japan** have remained almost the same for about 30 years.
<br>
<br>
```{r OECD Average Wages Data}
ggplot(subset(oecd_avgwage, country %in% picked_country), aes(year, avg_wage)) +
  geom_line(aes(color=country))
```

To explain the discrepancies between these countries, I decided to find out **what indicator is the most correlated to wage increases.**


## Alternative Indicator of Average Wages
Unfortunately, I was not able to find data on Average Wages on a global scale, so I will use **Nominal GDP** data as an alternative indicator (It is difficult for companies to increase wages when the country's economy is not growing. In other words, Average Wages tend to not grow when the GDP does not grow).

## Analysis
So, **what should a country do to increase its GDP and, in turn, its average wages?** 
<br>
<br>
Through analyzing many countries and indicators and their correlation to GDP growth, I concluded that **broad money growth has the strongest correlation to GDP growth.** 
<br>
<br>
Let me show you my analysis and visualizations below.

### Read a CSV file
```{r reading a file}
wb_gdp_n_broadmoney <- read_csv('/Users/hiranoseiji/Downloads/Data_Analysis/Macro_Economic_Wage_Analysis/worldbank_gdp_n_broadmoney.csv')
```
<br>
I manipulated the data using PostgreSQL and created a CSV file that contains data on GDP and broad money of over 150 countries. 
<br>
<br>
```{r sharing a file}
head(wb_gdp_n_broadmoney)
```
<br>
Below are **scatter plots showing correlation between GDP growth and broad money growth in Australia, United Kingdom, United States, and Japan** (data for German, France, and Italy was not available and I excluded data for Canada as it only had data until 2008).

### Australia
```{r visualizing AUS data}
ggscatter(subset(wb_gdp_n_broadmoney, code == "AUS"), x = "broad_money", y = "gdp", add = "reg.line") +
  stat_regline_equation(label.x = 100000000000, label.y = 1100000000000,  aes(label = ..rr.label..)) +
  stat_regline_equation(label.x = 100000000000, label.y = 1200000000000,  aes(label = ..eq.label..))
```

### United Kingdom
```{r visualizing GBR data}
ggscatter(subset(wb_gdp_n_broadmoney, code == "GBR"), x = "broad_money", y = "gdp", add = "reg.line") +
  stat_regline_equation(label.x = 250000000000, label.y = 2300000000000,  aes(label = ..rr.label..)) +
  stat_regline_equation(label.x = 250000000000, label.y = 2500000000000,  aes(label = ..eq.label..))
```

### United States
```{r visualizing USA data}
ggscatter(subset(wb_gdp_n_broadmoney, code == "USA"), x = "broad_money", y = "gdp", add = "reg.line") +
  stat_regline_equation(label.x = 1300000000000, label.y = 18000000000000,  aes(label = ..rr.label..)) +
  stat_regline_equation(label.x = 1300000000000, label.y = 20000000000000,  aes(label = ..eq.label..))
```

### Japan
```{r visualizing JPN data}
ggscatter(subset(wb_gdp_n_broadmoney, code == "JPN"), x = "broad_money", y = "gdp", add = "reg.line") +
  stat_regline_equation(label.x = 700000000000, label.y = 4400000000000,  aes(label = ..rr.label..)) +
  stat_regline_equation(label.x = 700000000000, label.y = 4800000000000,  aes(label = ..eq.label..))
```
<br>
<br>
It is very obvious that GDP growth and Broad Money growth is strongly correlated. **Circulating more money in the economy will help GDP growth which will then increase average wages.**
<br>
<br>
Here is **Year-over-Year line graphs on broad money** for the same countries.

### Australia
```{r AUS line chart}
ggplot(filter(wb_gdp_n_broadmoney, code == "AUS" & year %in% 1990:2020), aes(year, broad_money)) + geom_line()
```
```{r AUS bm growth rate}
round((filter(wb_gdp_n_broadmoney, code == "AUS" & year %in% 1990:2020)$broad_money[31] - filter(wb_gdp_n_broadmoney, code == "AUS" & year %in% 1990:2020)$broad_money[1])/filter(wb_gdp_n_broadmoney, code == "AUS" & year %in% 1990:2020)$broad_money[31]*100, digits = 2)
```
Broad Money has increased by **90.77%** since 1990 in Australia.

### United Kingdom
```{r GBR line chart}
ggplot(filter(wb_gdp_n_broadmoney, code == "GBR" & year %in% 1990:2020), aes(year, broad_money)) + geom_line()
```
```{r GBR bm growth rate}
round((filter(wb_gdp_n_broadmoney, code == "GBR" & year %in% 1990:2020)$broad_money[31] - filter(wb_gdp_n_broadmoney, code == "GBR" & year %in% 1990:2020)$broad_money[1])/filter(wb_gdp_n_broadmoney, code == "GBR" & year %in% 1990:2020)$broad_money[31]*100, digits = 2)
```
Broad Money has increased by **78.99%** since 1990 in the United Kingdom.

### United States
```{r USA line chart}
ggplot(filter(wb_gdp_n_broadmoney, code == "USA" & year %in% 1990:2020), aes(year, broad_money)) + geom_line()
```
```{r USA bm growth rate}
round((filter(wb_gdp_n_broadmoney, code == "USA" & year %in% 1990:2020)$broad_money[31] - filter(wb_gdp_n_broadmoney, code == "USA" & year %in% 1990:2020)$broad_money[1])/filter(wb_gdp_n_broadmoney, code == "USA" & year %in% 1990:2020)$broad_money[31]*100, digits = 2)
```
Broad Money has increased by **81.77%** since 1990 in the United States.

### Japan
```{r JPN line chart}
ggplot(filter(wb_gdp_n_broadmoney, code == "JPN" & year %in% 1990:2020), aes(year, broad_money)) + geom_line()
```
```{r JPN bm growth rate}
round((filter(wb_gdp_n_broadmoney, code == "JPN" & year %in% 1990:2020)$broad_money[31] - filter(wb_gdp_n_broadmoney, code == "JPN" & year %in% 1990:2020)$broad_money[1])/filter(wb_gdp_n_broadmoney, code == "JPN" & year %in% 1990:2020)$broad_money[31]*100, digits = 2)
```
Broad Money has increased by **58.96%** since 1990 in Japan.
<br>
<br>
<br>
This explains **why average wages stagnate in countries like Japan where broad money has not increased much in recent decades.** 
<br>
I also created a dashboard that contains data of other countries and published it on Tableau. If you are interested, click [here](https://public.tableau.com/app/profile/seiji.hirano/viz/MostCorrelatedIndicatortoCountrysAverageWagesGrowth/Dashboard1).
<br>

## Resources
[Average Wages Data from OECD](https://data.oecd.org/earnwage/average-wages.htm)<br>
[GDP Data from World Bank](https://data.worldbank.org/indicator/NY.GDP.MKTP.CD)<br>
[Broad Money Data from World Bank](https://data.worldbank.org/indicator/FM.LBL.BMNY.GD.ZS)
