---
title: "Die Roll Experiment Project"
author: "Seiji Hirano"
date: "2022-09-08"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About

This die roll experiment was introduced in a book written for beginner-level statistics learners by Yoichi Takahashi, a globally renowned Japanese economist who has served multiple prime ministers in policy-making. For more details about Yoichi Takahashi and his institute, click [here](https://www.highbridgeworld.com/).

# Explanation 
Mr. Takahashi used this die roll experiment example to explain about **histogram**. He says in the book that...

* If you roll one fair die three times, a combination of the three numbers has 216 outcomes (lowest:111, highest:666).
* This means that there is the probability of 0.00463 (1/216) of getting each combination.
* For this experiment, you number each combination from 1 through 216 (111 <- No.1, 112 <- No.2, ..., 666 <- No.216)
* Mr. Takahashi says that **if you repeat rolling a die and get a combination 30,000 times, all combinations will be picked with approximately the same frequency (around 139 times which is 30,000/216).**
* The key is that when all outcomes have an equal chance of occurring (discrete uniform distribution) like rolling a fair die, **the more samples you get, the smaller the variability of getting each outcome.**

As a beginner data analyst, I decided to conduct this experiment by myself with R and see if I can get the same result as explained in the book!




## Setting Up My Environment
Setting up my environment by loading the 'tidyverse' package.

```{r loading package}
library(tidyverse)
```



## Creating a Numbered-List-of-Combinations Table
### List of Combinations
I first created a table that contains **216 outcomes.**

```{r list of all combinations}
list.of.comb <- expand.grid(die1=c(1:6), die2=c(1:6),die3=c(1:6))

head(list.of.comb,10)
```

### Numbered List of Combinations
I then numbered each combination from 1 through 216.

```{r numbered list of all combinations}
numbered.list <- transmute(list.of.comb, num=seq(from=1,to=216,by=1), comb=as.integer(paste0(list.of.comb$die1,list.of.comb$die2,list.of.comb$die3)))

head(numbered.list)
```


## Creating a Die Roll Experiment Table
Here I created a 2nd table that contains **results after getting a combination 30,000 times.**

```{r Die Roll Experiment}
result <- numeric()

for (i in 1:30000) {
  die.roll <- sample(1:6, 3, replace = T)
  combined <- paste0(die.roll[1],die.roll[2],die.roll[3])
  char.to.int <- as.integer(combined)
  result <- c(result, char.to.int)
  df <- as.data.frame(result)
}

head(df)
```



## Joining the Two Tables
I joined numbered-list-of-combinations table & die-roll-experiment table

```{r joining the two tables}
joined.table <- merge(df,numbered.list,by.x = "result", by.y = "comb")

head(joined.table)
tail(joined.table)
```



## Final Table 
I ordered a joined table by 'num' and 'result' variables.

```{r Final Table}
final.table <- joined.table %>% 
  group_by(num,result) %>% 
  count()

head(final.table,10)
tail(final.table,10)

mean(final.table$n)
```

Looks like most outcomes are falling within **a range of 120-160 times.** 
Given that each combination has the probability of **0.00463 (1/216)** of being picked and that each combination should be picked **about 139 times (30,000 times 0.00463),** these results are likely correct.

Let's look at the visualizations to make sure.

## Visualization
### Histogram

```{r visualization}
pl <- ggplot(joined.table,aes(x=num))+
  geom_histogram(binwidth = 0.1)

print(pl)
```

### Density
```{r visualization part 2}
pl2 <- ggplot(final.table, aes(x=n)) +
  geom_density(fill="red", alpha=0.2) +
  geom_vline(xintercept = 139)

print(pl2)
```

* a vertical line shows the mean value of 139.


Let's try the entire experience few more times and see if each result shows similar visualizations.

### Attempt #2
```{r attempt 2}
result.2 <- numeric()

for (i in 1:30000) {
  die.roll.2 <- sample(1:6, 3, replace = T)
  combined.2 <- paste0(die.roll.2[1],die.roll.2[2],die.roll.2[3])
  char.to.int.2 <- as.integer(combined.2)
  result.2 <- c(result.2, char.to.int.2)
  df.2 <- as.data.frame(result.2)
}

joined.table.2 <- merge(df.2,numbered.list, by.x = "result.2", by.y = "comb")

final.table.2 <- joined.table.2 %>% 
  group_by(num,result.2) %>% 
  count() %>% 
  rename(n.2 = n)
```
```{r histogram n2}
pl.2 <- ggplot(joined.table.2,aes(x=num))+
  geom_histogram(binwidth = 0.1)
print(pl.2)
```
```{r density n2}
pl2.2 <- ggplot(final.table.2, aes(x=n.2)) +
  geom_density(fill="red", alpha=0.2) +
  geom_vline(xintercept = 139)
print(pl2.2)
```

### Attempt #3
```{r attempt 3}
result.3 <- numeric()

for (i in 1:30000) {
  die.roll.3 <- sample(1:6, 3, replace = T)
  combined.3 <- paste0(die.roll.3[1],die.roll.3[2],die.roll.3[3])
  char.to.int.3 <- as.integer(combined.3)
  result.3 <- c(result.3, char.to.int.3)
  df.3 <- as.data.frame(result.3)
}

joined.table.3 <- merge(df.3,numbered.list, by.x = "result.3", by.y = "comb")

final.table.3 <- joined.table.3 %>% 
  group_by(num,result.3) %>% 
  count() %>% 
  rename(n.3 = n)
```
```{r histogram n3}
pl.3 <- ggplot(joined.table.3,aes(x=num))+
  geom_histogram(binwidth = 0.1)
print(pl.3)
```
```{r density n3}
pl2.3 <- ggplot(final.table.3, aes(x=n.3)) +
  geom_density(fill="red", alpha=0.2) +
  geom_vline(xintercept = 139)
print(pl2.3)
```

### Attempt #4
```{r attempt 4}
result.4 <- numeric()

for (i in 1:30000) {
  die.roll.4 <- sample(1:6, 3, replace = T)
  combined.4 <- paste0(die.roll.4[1],die.roll.4[2],die.roll.4[3])
  char.to.int.4 <- as.integer(combined.4)
  result.4 <- c(result.4, char.to.int.4)
  df.4 <- as.data.frame(result.4)
}

joined.table.4 <- merge(df.4,numbered.list, by.x = "result.4", by.y = "comb")

final.table.4 <- joined.table.4 %>% 
  group_by(num,result.4) %>% 
  count() %>% 
  rename(n.4 = n)
```
```{r histogram n4}
pl.4 <- ggplot(joined.table.4,aes(x=num))+
  geom_histogram(binwidth = 0.1)
print(pl.4)
```
```{r density n4}
pl2.4 <- ggplot(final.table.4, aes(x=n.4)) +
  geom_density(fill="red", alpha=0.2) +
  geom_vline(xintercept = 139)
print(pl2.4)
```

### Attempt #5
```{r attempt 5}
result.5 <- numeric()

for (i in 1:30000) {
  die.roll.5 <- sample(1:6, 3, replace = T)
  combined.5 <- paste0(die.roll.5[1],die.roll.5[2],die.roll.5[3])
  char.to.int.5 <- as.integer(combined.5)
  result.5 <- c(result.5, char.to.int.5)
  df.5 <- as.data.frame(result.5)
}

joined.table.5 <- merge(df.5,numbered.list, by.x = "result.5", by.y = "comb")

final.table.5 <- joined.table.5 %>% 
  group_by(num,result.5) %>% 
  count() %>% 
  rename(n.5 = n)
```
```{r histogram n5}
pl.5 <- ggplot(joined.table.5,aes(x=num))+
  geom_histogram(binwidth = 0.1)
print(pl.5)
```
```{r density n5}
pl2.5 <- ggplot(final.table.5, aes(x=n.5)) +
  geom_density(fill="red", alpha=0.2) +
  geom_vline(xintercept = 139)
print(pl2.5)
```

## Conclusion

**From my die roll experiment, I was able to get the result that backs up Mr. Takahashi's explanation in the book!**
